PWD.KAZ

### 2019-07-19 results/PWD.KAZ.BivData.Rmd
clean up of PWD.KAZ analysis script and associated files

#curBivData_18may19_PWD_m1 = read.csv("~./MLH1repo/results/curate_18may18_PWD_m1_BivData.txt", header = TRUE, sep = '\t')
#curBivData_21may18_KAZ_m2 = read.csv("~./MLH1repo/results/curate_21may18_KAZ_m2_BivData.txt",header = TRUE, sep = '\t')

#PWD set 4 mice, ~100 objs each)
PWD_curate_BivData.full =
  read.csv("~./MLH1repo/data/BivData/Merged_Curate_PWD_BivData.csv", header = TRUE)
#remove bad segmentation
PWD_curate_BivData <- PWD_curate_BivData.full[PWD_curate_BivData.full$SC.pass == 1,]

@@

#load hand measures, get it to match the auto
PWD_hand_measure_BivData.full =
  read.csv("~./MLH1repo/data/BivData/PWD_male_hand_measures.csv", header = TRUE)

#remove extra empty rows
PWD_hand_measure_BivData.full <- PWD_hand_measure_BivData.full[!(is.na(PWD_hand_measure_BivData.full$fileName) | PWD_hand_measure_BivData.full$fileName == ""), ]

#PWD_hand_measure_BivData.full$Foci1  <- PWD_hand_measure_BivData.full$foci.1.position
#PWD_hand_measure_BivData.full$Foci2 <- PWD_hand_measure_BivData.full$medial.foci.position
#PWD_hand_measure_BivData.full$Foci3 <- PWD_hand_measure_BivData.full$distal.foci.position
#PWD_hand_measure_BivData.full$boxNumber <- PWD_hand_measure_BivData.full$manual.blobject.name

PWD_hand_measure_BivData.full$Obj.ID <- paste(PWD_hand_measure_BivData.full$fileName, PWD_hand_measure_BivData.full$boxNumber, sep = "_")

PWD_hand_measure_BivData.full$centromere_PER_Position <- PWD_hand_measure_BivData.full$centromere_ABS_Position / PWD_hand_measure_BivData.full$chromosomeLength

PWD_hand_measure_BivData.full$SC.pass <- "1"
PWD_hand_measure_BivData.full$foci.pass <- "1"
PWD_hand_measure_BivData.full$curated.1pass.0fail. <- ""
#PWD_hand_measure_BivData.full$curation.notes <- ""
#PWD_hand_measure_BivData.full$hand.foci.count <- ""

source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
PWD_hand_measure_BivData.full <- add_mouse(PWD_hand_measure_BivData.full)
PWD_hand_measure_BivData.full <- add_category(PWD_hand_measure_BivData.full)
PWD_hand_measure_BivData.full <- add_strain(PWD_hand_measure_BivData.full)
PWD_hand_measure_BivData.full <- add_sex(PWD_hand_measure_BivData.full)

#3 columns don't overlap
#avergeRED, numberCrossOvers, 
#manual.blobject.name,

KAZ_hand_measure_BivData.full =
  read.csv("~./MLH1repo/data/BivData/KAZ_male_hand_measures.csv", header = TRUE)

KAZ_hand_measure_BivData.full <- KAZ_hand_measure_BivData.full[!(is.na(KAZ_hand_measure_BivData.full$fileName) | KAZ_hand_measure_BivData.full$fileName == ""), ]

KAZ_hand_measure_BivData.full$boxNumber <- KAZ_hand_measure_BivData.full$manual.blobject.name

KAZ_hand_measure_BivData.full$Obj.ID <- paste(KAZ_hand_measure_BivData.full$fileName, KAZ_hand_measure_BivData.full$boxNumber, sep = "_")

KAZ_hand_measure_BivData.full$SC.pass <- "1"
KAZ_hand_measure_BivData.full$foci.pass <- "1"
KAZ_hand_measure_BivData.full$curated.1pass.0fail. <- ""

source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
KAZ_hand_measure_BivData.full <- add_mouse(KAZ_hand_measure_BivData.full)
KAZ_hand_measure_BivData.full <- add_category(KAZ_hand_measure_BivData.full)
KAZ_hand_measure_BivData.full <- add_strain(KAZ_hand_measure_BivData.full)
KAZ_hand_measure_BivData.full <- add_sex(KAZ_hand_measure_BivData.full)

#not sure why I have a distinction of hand_measure and hand_measure.full
#KAZ_hand_measure_BivData <- KAZ_hand_measure_BivData.full[KAZ_hand_measure_BivData.full$SC.pass == 1,]

#manual blobject name -- boxNumber, Obj.ID 
#  20jul15_18may15_PWD_m1_sp1_2.2_blob1
```

@@

### BivData dataset

Goal:  Count the Bivdata for pwd and KAZ males
start determining the subset of data

```{r, echo=FALSE}
#Curated data
#merge PWD and KAZ BivData

#extract just the relavant PWD and KAZ
PWD.MLH1 <- MLH1_data[MLH1_data$category == "PWD male",]
KAZ.MLH1 <- MLH1_data[MLH1_data$category == "KAZ male",]

PWD.BivData.cells <- BivData_cells[BivData_cells$category == "PWD male",]
PWD.BivData <- fullBivData[fullBivData$category == "PWD male",]

KAZ.BivData.cells <- BivData_cells[BivData_cells$category == "KAZ male",]
KAZ.BivData <- fullBivData[fullBivData$category == "KAZ male",]

PWD.KAZ_MLH1<- rbind(PWD.MLH1, KAZ.MLH1)
PWD.KAZ_BivData<- rbind(PWD.BivData, KAZ.BivData)

what are the distributions of SC lengths?

```{r cell.level.tables, echo=FALSE}

#make cell level tables
#id the 18-20 biv cells

#estimate which cells can be used to make grouped pools

@@

```{r}
#SC lengths by CO number

foci.pass <- KAZ_curate_BivData

@@

```{r SClength.cell.tables}
#add total COs
KAZ_BivData_cells <- ddply(KAZ_curate_BivData, c("fileName"), summarise,

cells2curate <- KAZ_BivData_cells[KAZ_BivData_cells$pass_rate > .7,]
#write.table(cells2curate, "~./MLH1repo/results/KAZ.cell.curate.csv", sep=",", row.names = FALSE)

```{r PWD.merge.manual.auto, echo=FALSE}
#merging manual measures and auto measures
#KAZ_hand_measure_BivData
#PWD_hand_measure_BivData

#auto.curate
auto.c_names <- c("fileName","centromere_ABS_Position","centromere_PER_Position",  "chromosomeLength","boxNumber","Obj.ID","SC.pass","foci.pass","curated.1pass.0fail.","curation.notes","hand.foci.count","Foci1","Foci2","Foci3" ,"mouse", "strain","sex","category")
new.DF <- PWD_curate_BivData[ , auto.c_names, drop = TRUE]

#manual
manual_names <- c("fileName","chromosomeLength","centromere_ABS_Position","boxNumber","Obj.ID","Foci1","Foci2","Foci3","centromere_PER_Position","SC.pass","foci.pass","curated.1pass.0fail.", "mouse", "category","strain","sex","curation.notes", "hand.foci.count")
new.DF2 <- PWD_hand_measure_BivData.full[ , manual_names, drop = TRUE]

#2. merge the 2 DFs
PWD.auto.man_BivData <- rbind(new.DF, new.DF2)

#some cleanup may be required for the manual data; remove all XY, mark chrm classes for all 
#make better variable name
PWD.auto.man_BivData_cells <- ddply(PWD.auto.man_BivData, c("fileName"), summarise,
                    number = length(Obj.ID),
                    #make the rank or
                     maxChrm_length = max(chromosomeLength),
                    maxChrm_boxNumber = boxNumber[which.max(chromosomeLength)]

#Isolate cells with all autosomes measured
PWDfull.cell.list <- PWD.auto.man_BivData_cells[PWD.auto.man_BivData_cells$number > 17,]
PWDfull.cell.list <- PWDfull.cell.list[PWDfull.cell.list$number < 21,]#remove the too many cell

#this gives the rank to each biv within each cell  
PWD.Bivs.w.rank <- ddply(PWD.auto.man_BivData, .(fileName), transform, rank = rank(chromosomeLength,ties.method ="min")-1)
#there are instances of ties,

#this does something weird
#PWD.auto.man_BivData$rank1 <- ddply(PWD.auto.man_BivData, .(fileName), transform, rank = rank(chromosomeLength,ties.method ="min")-1)

#pull all biv data for the full cell data
PWD_whole.cell.BivData <- PWD.Bivs.w.rank[PWD.Bivs.w.rank$fileName %in% PWDfull.cell.list$fileName,]
#ok these should be full cells

#there are 2x as many 0's bc 1 and 0 are binned together
#update all hand.foci.counts
#y <- droplevels(y)

bin.short.PWD <- PWD_whole.cell.BivData[PWD_whole.cell.BivData$rank < 5,]

bin.short.PWD <- bin.short.PWD[bin.short.PWD$fileName %in% full.cell.list$fileName,]

#short bin patterns vs long bins / total?
#eventually I want to compare NORM FOCI POSITIONS patterns of these bins across PWD and KAZ
#merging is a little too confusing... 
#just parse down and do rbind?
m1 <- merge(PWD_curate_BivData, PWD_hand_measure_BivData.full, by.x = Obj.ID)

#KAZ_curate_BivData
#KAZ_hand_measure_BivData.full

hist(hh2$rank, breaks =21)
```

@@

```{r KAZ.manual.merge}
#KAZ_hand_measure_BivData
#PWD_hand_measure_BivData
#KAZ_curate_BivData <- KAZ_curate_BivData.full[KAZ_curate_BivData.full$SC.pass == 1,]

#1 match the colnames,  #merge, partial match
colnames(KAZ_curate_BivData)
colnames(KAZ_hand_measure_BivData.full)

KAZ_hand_measure_BivData.full$centromere_PER_Position <- KAZ_hand_measure_BivData.full$centromere_ABS_Position / KAZ_hand_measure_BivData.full$chromosomeLength

KAZ_hand_measure_BivData.full$curation.notes <- KAZ_hand_measure_BivData.full$notes

#auto; avergeRED, numberCrossOvers, 20
#manual; 15

#auto.curate
KAZ.auto.c_names <- c("fileName","centromere_ABS_Position","centromere_PER_Position",  "chromosomeLength","boxNumber","Obj.ID","SC.pass","foci.pass","curated.1pass.0fail.","curation.notes","hand.foci.count","Foci1","Foci2","Foci3" ,"mouse", "strain","sex","category")
new.DF <- KAZ_curate_BivData[ , KAZ.auto.c_names, drop = TRUE]

#manual
KAZ.manual_names <- c("fileName","chromosomeLength","centromere_ABS_Position","boxNumber","Obj.ID","Foci1","Foci2","Foci3","centromere_PER_Position", "SC.pass","foci.pass","curated.1pass.0fail.", "mouse", "category","strain","sex","curation.notes", "hand.foci.count")
new.DF2 <- KAZ_hand_measure_BivData.full[ , KAZ.manual_names, drop = TRUE]

#2. merge the 2 DFs
KAZ.auto.man_BivData <- rbind(new.DF, new.DF2)

#some cleanup may be required for the manual data; remove all XY, mark chrm classes for all 
#make better variable name
KAZ.auto.man_BivData_cells <- ddply(KAZ.auto.man_BivData, c("fileName"), summarise,
                    number = length(Obj.ID),
                    #make the rank or
                     maxChrm_length = max(chromosomeLength),
                    maxChrm_boxNumber = boxNumber[which.max(chromosomeLength)]

#Isolate cells with all autosomes measured
KAZfull.cell.list <- KAZ.auto.man_BivData_cells[KAZ.auto.man_BivData_cells$number > 17,]
KAZfull.cell.list <- KAZfull.cell.list[KAZfull.cell.list$number < 21,]#remove the too many cell

#this gives the rank to each biv within each cell  
KAZ.Bivs.w.rank <- ddply(KAZ.auto.man_BivData, .(fileName), transform, rank = rank(chromosomeLength,ties.method ="min")-1)
#there are instances of ties,

KAZ_whole.cell.BivData <- KAZ.Bivs.w.rank[KAZ.Bivs.w.rank$fileName %in% KAZfull.cell.list$fileName,]
#ok these should be full cells
hist(KAZ_whole.cell.BivData$rank, breaks =25)
KAZ_whole.cell.BivData <- droplevels(KAZ_whole.cell.BivData)
table(KAZ_whole.cell.BivData$fileName, KAZ_whole.cell.BivData$rank)

@@

```
##########################
# MELTED CHRM1 POSITIONS #
##########################

```{r PWD.melt.foci.data, echo=FALSE}
library(reshape2)

#newPWD_curate_BivData =
#  read.csv("~./MLH1repo/data/BivData/Merged_Curate_PWD_BivData.csv", header = TRUE)

#update these dataframes!
#isolate the passing foci 
#curBivData_PWD.good.foci  <- newPWD_curate_BivData[newPWD_curate_BivData$foci.pass == 1,]

curBivData_PWD.good.foci <- PWD.auto.man_BivData[PWD.auto.man_BivData$foci.pass == 1,]

#create normalized foci values!
curBivData_PWD.good.foci$nrm.F1 <- curBivData_PWD.good.foci$Foci1 / curBivData_PWD.good.foci$chromosomeLength
curBivData_PWD.good.foci$nrm.F2 <- curBivData_PWD.good.foci$Foci2 / curBivData_PWD.good.foci$chromosomeLength
curBivData_PWD.good.foci$nrm.F3 <- curBivData_PWD.good.foci$Foci3 / curBivData_PWD.good.foci$chromosomeLength

#melting the DataFrame is needed to ploting the normalized foci positions

#do not include cols for IFD, or ABS.foci
curBivData_PWD.melt <- melt( curBivData_PWD.good.foci[,c("fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
    "hand.foci.count","nrm.F1","nrm.F2","nrm.F3",  "mouse", "category",   "strain" , "sex")] ,

colnames(curBivData_PWD.melt) <- c( "fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
    "hand.foci.count", "mouse", "category",   "strain" , "sex",
                                       #name for the 2 melted cols
                                      "Foci.Type.nrm", "Foci.nrm.pos")

#remove NAs
curBivData_PWD.melt <- curBivData_PWD.melt[!(is.na(curBivData_PWD.melt$Foci.nrm.pos) | curBivData_PWD.melt$Foci.nrm.pos == ""), ]

#table(newPWD_curate_BivData$fileName)
#table(newPWD_curate_BivData$mouse) #4 mice, ~100 each mouse

```{r melt.KAZ, echo=FALSE}

#use the combined - auto and manual data

curBivData_KAZ.good.foci <- KAZ.auto.man_BivData[KAZ.auto.man_BivData$foci.pass == 1,]
#curated objects
#KAZ_m2_curate_bivData =
#  read.csv("~./MLH1repo/data/BivData/curated_21may18_KAZ_m2_BivData.csv", header = TRUE)

KAZ_m2_curate_bivData$nrm.F1.pos <- KAZ_m2_curate_bivData$Foci1.pos / KAZ_m2_curate_bivData$chromosomeLength
KAZ_m2_curate_bivData$nrm.F2.pos <- KAZ_m2_curate_bivData$Foci2.pos / KAZ_m2_curate_bivData$chromosomeLength
KAZ_m2_curate_bivData$nrm.F3.pos <-   KAZ_m2_curate_bivData$Foci3.pos / KAZ_m2_curate_bivData$chromosomeLength

#remove those that don't pass
KAZ_m2_curate_bivData_SC_pass <- KAZ_m2_curate_bivData[KAZ_m2_curate_bivData$SC.pass == 1,]
#what else...

table(KAZ_m2_curate_bivData_SC_pass$fileName)

#melt the foci positions, best thing to compare for cells that don't have all the bivalents
#ggplots

```{r meltKAZ, echo=FALSE}

#melt KAZ
#isolate the passing foci 
#curBivData_KAZ.good.foci  <- curBivData_KAZ[curBivData_KAZ$foci.pass == 1,]

curBivData_KAZ.good.foci <- KAZ.auto.man_BivData[KAZ.auto.man_BivData$foci.pass == 1,]
colnames(curBivData_KAZ.good.foci)

curBivData_KAZ.good.foci$nrm.F1.pos <- curBivData_KAZ.good.foci$Foci1 / curBivData_KAZ.good.foci$chromosomeLength
curBivData_KAZ.good.foci$nrm.F2.pos <- curBivData_KAZ.good.foci$Foci2 / curBivData_KAZ.good.foci$chromosomeLength
curBivData_KAZ.good.foci$nrm.F3.pos <-   curBivData_KAZ.good.foci$Foci3 / curBivData_KAZ.good.foci$chromosomeLength

#melting the DataFrame is needed to ploting the normalized foci positions

#do not include cols for IFD, or ABS.foci
curBivData_KAZ.melt <- melt( curBivData_KAZ.good.foci[,c("fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
    "hand.foci.count","nrm.F1.pos","nrm.F2.pos","nrm.F3.pos",  "mouse", "category",   "strain" , "sex")] ,

colnames(curBivData_KAZ.melt) <- c( "fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
    "hand.foci.count", "mouse", "category",   "strain" , "sex",
                                       #name for the 2 melted cols
                                      "nrmFoci.Type", "nrmFoci.pos")

#remove NAs
curBivData_KAZ.melt <- curBivData_KAZ.melt[!(is.na(curBivData_KAZ.melt$nrmFoci.Type) | curBivData_KAZ.melt$nrmFoci.pos==""), ]

@@

```{r curatedBivData}

curBivData_KAZ = read.csv("~./MLH1repo/data/BivData/curate_KAZ_BivData.csv", header = TRUE, sep = ',')
#curBivData_PWD = read.csv("~./MLH1repo/data/BivData/curate_PWD_BivData.csv", header = TRUE, sep = ',')

#make normalized

curBivData_KAZ$nrm.F1.pos <- curBivData_KAZ$Foci1 / curBivData_KAZ$chromosomeLength
curBivData_KAZ$nrm.F2.pos <- curBivData_KAZ$Foci2 / curBivData_KAZ$chromosomeLength
curBivData_KAZ$nrm.F3.pos <-   curBivData_KAZ$Foci3 / curBivData_KAZ$chromosomeLength

curBivData_PWD$nrm.F1.pos <- curBivData_PWD$Foci1 / curBivData_PWD$chromosomeLength
curBivData_PWD$nrm.F2.pos <- curBivData_PWD$Foci2 / curBivData_PWD$chromosomeLength
curBivData_PWD$nrm.F3.pos <-   curBivData_PWD$Foci3 / curBivData_PWD$chromosomeLength

#table / what are the numbers of observations
table(curBivData_PWD$fileName)
#biv measures per cell

SC.passed.PWD <- curBivData_PWD[curBivData_PWD$SC.pass == 1,]#299 -> 208
table(SC.passed.PWD$fileName)

SC.passed.KAZ <- curBivData_KAZ[curBivData_KAZ$SC.pass == 1,]#299 -> 208

@@

```{r fit.exp}
#fitting IFDs to expoential Distribution
#curBivData_PWD.melt
#make an IFD plot of Null.IFDs drawn from exponential
#curBivData_KAZ.melt2

@@

fit.PWD <- fitdistr(all.2CO$IFD1_PER[all.2CO$strain == "PWD"], "exponential")#change to gamma
fit.KAZ <- fitdistr(all.2CO$IFD1_PER[all.2CO$strain== "KAZ"], "exponential")

PWD.curve <- hist(all.2CO$IFD1_PER[all.2CO$strain == "PWD"], freq = FALSE, breaks = 100, xlim = c(0, quantile(ex, 0.99)))+
curve(dexp(x, rate = fit.PWD$estimate), from = 0, col = "red", add = TRUE)

@@

```{r rand.PWD, echo=FALSE}

PWD.BivData <- fullBivData[fullBivData$category == "PWD male",]#670
cur18may15_PWD_m1 <- PWD.BivData.curate.list[PWD.BivData.curate.list$mouse == "18may15_PWD_m1",]
cur18may15_PWD_m2 <- PWD.BivData.curate.list[PWD.BivData.curate.list$mouse == "18may15_PWD_m2",]

cur10mar15_PWD_m1 <- PWD.BivData.curate.list[PWD.BivData.curate.list$mouse == "10mar15_PWD_m1",]
cur10mar15_PWD_m2 <- PWD.BivData.curate.list[PWD.BivData.curate.list$mouse == "10mar15_PWD_m2",]

table(PWD.BivData.curate.list$mouse)
#4mice in the current dataset

```{r KAZ.curate.list, echo=FALSE}
#write out list for initial quality control and subsetting

#remove centromere_PER > .2

KAZ.Biv.curate.list <- KAZ_bivData[KAZ_bivData$centromere_PER_Position < 0.2,]#3123

@@

#KAZ_mice

#11jan18_KAZ_m2-curated file
#11jan18_KAZ_m3-
#11jan18_KAZ_m4-

@@

>>>>

#1. Curated Data
#PWD set 4 mice, ~100 objs each)
PWD_curate_BivData.full =
  read.csv("~./MLH1repo/data/BivData/Merged_Curate_PWD_BivData.csv", header = TRUE)
#remove bad segmentation
PWD_curate_BivData <- PWD_curate_BivData.full[PWD_curate_BivData.full$SC.pass == 1,]

@@

#merged KAZ and PWD
curate_BivData <- rbind(KAZ_curate_BivData, PWD_curate_BivData)

#2. Hand measures (merge first, then add extra cols)
PWD_hand_measure_BivData =
  read.csv("~./MLH1repo/data/BivData/PWD_male_hand_measures.csv", header = TRUE)

KAZ_hand_measure_BivData =
  read.csv("~./MLH1repo/data/BivData/KAZ_male_hand_measures.csv", header = TRUE)

#merge together -- check that the colnames match!!
HandMeasures_BivData <- rbind(KAZ_hand_measure_BivData, PWD_hand_measure_BivData)

#remove extra empty rows
HandMeasures_BivData <- HandMeasures_BivData[!(is.na(HandMeasures_BivData$fileName) | HandMeasures_BivData$fileName == ""), ]

#add Obj.ID
HandMeasures_BivData$Obj.ID <- paste(HandMeasures_BivData$fileName, HandMeasures_BivData$boxNumber, sep = "_")
#add centromere_PER (match auto)
HandMeasures_BivData$centromere_PER_Position <- HandMeasures_BivData$centromere_ABS_Position / HandMeasures_BivData$chromosomeLength

#more columns to match Auto DFs
HandMeasures_BivData$SC.pass <- "1"
HandMeasures_BivData$foci.pass <- "1"
HandMeasures_BivData$curated.1pass.0fail. <- ""

source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
HandMeasures_BivData <- add_mouse(HandMeasures_BivData)
HandMeasures_BivData <- add_category(HandMeasures_BivData)
HandMeasures_BivData <- add_strain(HandMeasures_BivData)
HandMeasures_BivData <- add_sex(HandMeasures_BivData)

#make indevidal DFs by strain just in case...
PWD.BivData.Hand_measures <- HandMeasures_BivData[HandMeasures_BivData$strain == "PWD",]
KAZ.BivData.Hand_measures <- HandMeasures_BivData[HandMeasures_BivData$strain == "KAZ",]

```{r Merge.manual.auto, echo=FALSE}

#merge with the both strain DF
#curate_BivData ... HandMeasures_BivData

#auto.curate
auto.c_names <- c("fileName","centromere_ABS_Position","centromere_PER_Position",  "chromosomeLength","boxNumber","Obj.ID","SC.pass","foci.pass","curated.1pass.0fail.","curation.notes","hand.foci.count","Foci1","Foci2","Foci3" ,"mouse", "strain","sex","category")
new.DF <- curate_BivData[ , auto.c_names, drop = TRUE]

#manual
manual_names <- c("fileName","chromosomeLength","centromere_ABS_Position","boxNumber","Obj.ID","Foci1","Foci2","Foci3","centromere_PER_Position","SC.pass","foci.pass","curated.1pass.0fail.", "mouse", "category","strain","sex","curation.notes", "hand.foci.count")
new.DF2 <- HandMeasures_BivData[ , manual_names, drop = TRUE]

#Merge the 2 DFs
Auto.Man_BivData <- rbind(new.DF, new.DF2)

#make WHOLE CELL DF

#some cleanup may be required for the manual data; remove all XY, mark chrm classes for all 
#make better variable name
Auto.Man_BivData_cells <- ddply(Auto.Man_BivData, c("fileName"), summarise,
                    number = length(Obj.ID),
                    #make the rank or
                     maxChrm_length = max(chromosomeLength),
                    maxChrm_boxNumber = boxNumber[which.max(chromosomeLength)]

#Isolate cells with all (19) autosomes measured
whole.cell_table <- Auto.Man_BivData_cells[Auto.Man_BivData_cells$number > 17,]
whole.cell_table <- whole.cell_table[whole.cell_table$number < 21,]
whole.cell.list <- unique(whole.cell.DF$fileName)

#make the whole.cell BivData DF
whole.cell.DF <- Auto.Man_BivData[Auto.Man_BivData$fileName %in% whole.cell.list,]

#ADD within cell RANK to indiviual Bivalents,
whole.cell.DF <- ddply(whole.cell.DF, .(fileName), transform, rank = rank(chromosomeLength,ties.method ="min")-1)

#there are instances of ties, 
#oo -- look at correlation ranks

#make the bins
short.bin.Bivs <- whole.cell.DF[whole.cell.DF$rank < 6,]
long.bin.Bivs <- whole.cell.DF[whole.cell.DF$rank > 14,]

```{r Melt.DF}
library(reshape2)
#Auto.Man_BivData, all curate data
#get the whole cell verion with whole.cell.list

#isolate bivs with passing foci measures
Auto.Man_BivData.good.foci <- Auto.Man_BivData[Auto.Man_BivData$foci.pass == 1,]
colnames(Auto.Man_BivData.good.foci)

#add nrm foci measures
Auto.Man_BivData.good.foci$nrm.F1.pos <- Auto.Man_BivData.good.foci$Foci1 / Auto.Man_BivData.good.foci$chromosomeLength
Auto.Man_BivData.good.foci$nrm.F2.pos <- Auto.Man_BivData.good.foci$Foci2 / Auto.Man_BivData.good.foci$chromosomeLength
Auto.Man_BivData.good.foci$nrm.F3.pos <-   Auto.Man_BivData.good.foci$Foci3 / Auto.Man_BivData.good.foci$chromosomeLength

#melting the DataFrame is needed to ploting the normalized foci positions
#do not include cols for IFD, or ABS.foci
Auto.Man_BivData.melt <- melt( Auto.Man_BivData.good.foci[,c("fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
    "hand.foci.count","nrm.F1.pos","nrm.F2.pos","nrm.F3.pos",  "mouse", "category",   "strain" , "sex")] ,

colnames(Auto.Man_BivData.melt) <- c( "fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
    "hand.foci.count", "mouse", "category",   "strain" , "sex",
                                       #name for the 2 melted cols
                                      "nrmFoci.Type", "nrmFoci.pos")

#remove NAs
Auto.Man_BivData.melt <- Auto.Man_BivData.melt[!(is.na(Auto.Man_BivData.melt$nrmFoci.Type) | Auto.Man_BivData.melt$nrmFoci.pos==""), ]

#make PWD and KAZ specific melt.DF ... if you want

#make whole cell melt
whole.cell_melt.DF <- Auto.Man_BivData.melt[Auto.Man_BivData.melt$fileName %in% whole.cell.list,]
#add rank to these measures
whole.cell_melt.DF <- ddply(whole.cell_melt.DF, .(fileName), transform, rank = rank(chromosomeLength,ties.method ="min")-1)

```{r saveRdata}
# all the main DFs have been made ...
#save rdata file!

save.image("data/MLH1/MLH1_data_setup_62419.RData")
```

```{r SClength.cell.tables}
#this is useful for making the list of cells to curate (comment out after curation step)

#add total COs
KAZ_BivData_cells <- ddply(KAZ_curate_BivData, c("fileName"), summarise,

#cells2curate <- KAZ_BivData_cells[KAZ_BivData_cells$pass_rate > .7,]
#write.table(cells2curate, "~./MLH1repo/results/KAZ.cell.curate.csv", sep=",", row.names = FALSE)

Data frames and analyses are the main output of this script - for a saved Rdata file.

1. Curated BivData; read files, merge/rbind together

2. Hand Measures; read files, rbind together

3. Auto+manual; combine the curate and hand measures, rbind both strains together.
    3B. Auto-Manual_whole_cell, make a list of fileNames for extracting these measures later; the biv from cells with all 19 autosome measures. Add bivalent rank to this DF.

4. melted DF; Melt the Foci-Pass of Auto_Manual_DF

--- Once the above DF's are made, proceed with following analyses ---

1. SC Length
    a. raw / loose bivalents
    b. binned bivalents, short vs long, chrm class

2. Normalized Foci Position
   a. all/full
   b. bins

3. 2-CO
    a. IFDs, raw norm
    b. F1 by F2

4. Rbar?

@@

PWD and KAZ differ in gwRR/MLH1 counts, is there a correlation in meiotic features which go with expected differences in gwRR.

@@

```{r}

@@

@@

@@

@@

@@

```{r fit.exp, echo=FALSE}
#fitting IFDs to expoential Distribution
#curBivData_PWD.melt
#make an IFD plot of Null.IFDs drawn from exponential
#curBivData_KAZ.melt2

@@

fit.PWD <- fitdistr(all.2CO$IFD1_ABS[all.2CO$strain == "PWD"], "gamma")#change to gamma, (1 outlier)

fit.KAZ <- fitdistr(all.2CO$IFD1_ABS[all.2CO$strain== "KAZ"], "gamma")

PWD.curve <- hist(all.2CO$IFD1_ABS[all.2CO$strain == "PWD"], freq = FALSE, breaks = 100, xlim = c(0, quantile(ex, 0.99)))+
curve(dexp(x, rate = fit.PWD$estimate), from = 0, col = "red", add = TRUE)

@@

```{r makeKAZ.curate.list, echo=FALSE}
#write out list for initial quality control and subsetting
#remove centromere_PER > .2
KAZ.Biv.curate.list <- KAZ_bivData[KAZ_bivData$centromere_PER_Position < 0.2,]#3123

@@

@@

#y <- droplevels(y)
